### 2018-04-24 contents/reviews/pain.tex
Updates from ShareLaTeX

An important contribution of the gate-control theory was to emphasize central nervous system mechanisms \cite{Melzack1999}, and the role of the dorsal horn in pain processing, and therefore reconciles the evidences of pain-specific primary afferent and with pattern theory. Contemporary views consider pain to be complex and multi-dimensional, where the processing of ascending nociceptive signals, in conjunction with descending modulatory pathways, shape the pain experience. Melzack and Casey further proposed that pain consists of sensory-discriminative, affective-motivational, and cognitive-evaluative components \cite{Melzack1968}. Specifically, the sensory-discriminative component consists of peripheral nociceptive afferents transmit to the spinal dorsal horn (laminae V), whereby nociceptive specific (NS) and wide-dynamic-range (WDR) neurons convergently process these primary afferents, and relay them to the thalamus, and subsequently the primary sensory cortex. A central concept, which was also introduced by Melzack, is the  neuromatrix theory of pain, which dictates that the pain experience is produced by a "neurosignature" pattern, that is generated from widely distributed neural network, which Melzack refers as the "body-self neuromatrix". \cite{Melzack1999}. The concept is now more commonly referred to as the "pain matrix" \cite{tracey2007a}. (See section \ref{section:neuromatrix})

@@

 Neuropathic and acute nociceptive pain have a high degree of overlap in brain activations \cite{Moisset2007}, therefore differentiating between neuropathic and acute pain from neuroimaging continue to be an active area of research. Brain regions that are involved in the somatosensation of pain process are involved in both acute and chronic pain. These regions include the posterior insula, lateral thalamus and dorsal striatum. In chronic pain conditions however, there is increased activation in the anterior insula, ventral striatum, orbital frontal cortex and dorsal lateral prefrontal cortex \cite{Lorenz2005}. In chronic back pain, diffusion changes in fractional anisotropy was correlated with medial prefrontal cortex and nucleus accumbens connectivity in pain recovery, but not when the pain persistented \cite{Mansour2013}. Evidence also suggest that back pain chronification seem to increase activation in the emotional regions (amygada and medial prefrontal cortex) of the brain, and transition away from sensory region activations \cite{Hashmi2013}. A shift in forebrain activities is observed in neuropathic versus acute pain, abnormal activities in anterior cingulate cortex and prefrontal cortex is commonly observed in these studies \cite{Davis2006}.

 Advances in modern machine learning methods led to their uses in attempting to improve neuropathic pain classification and diagnosis. Modern approaches combines cross-modal and additional analytical features, and can involve fMRI, diffusion imaging, as well as functional and/or structural connectivity. Machine learning in the form of multivariate pattern analysis to classify individuals with chronic back pain from normal controls, and achieve 92\% accuracy \cite{Callan2014}. Other studies in chronic pelvic pain using linear support vector machine was able to achieve 73\% accuracy \cite{Bagarinao2014}. Differentiating neuropathic and acute pain however is more difficult. Functional connectivity based classification on differentiating fibromyalgia, rheumatoid arthritis and controls was able to achieve a diagnositc accuracy of 69\%, and classification accuracy of 78\% \cite{Sundermann2014}.

>>>>

An important contribution of the gate-control theory was to emphasize central nervous system mechanisms \cite{Melzack1999}, and the role of the dorsal horn in pain processing, and therefore reconciles the evidences of pain-specific primary afferent and with pattern theory. Contemporary views consider pain to be complex and multi-dimensional, where the processing of ascending nociceptive signals, in conjunction with descending modulatory pathways, shape the pain experience. Melzack and Casey further proposed that pain consists of sensory-discriminative, affective-motivational, and cognitive-evaluative components \cite{Melzack1968}. Specifically, the sensory-discriminative component consists of peripheral nociceptive afferents transmit to the spinal dorsal horn (laminae V), whereby nociceptive specific (NS) and wide-dynamic-range (WDR) neurons convergently process these primary afferents, and relay them to the thalamus, and subsequently the primary sensory cortex. A central concept, which was also introduced by Melzack, is the  neuromatrix theory of pain, which dictates that the pain experience is produced by a "neurosignature" pattern, that is generated from widely distributed neural network, which Melzack refers as the "body-self neuromatrix". \cite{Melzack1999}. The concept is now more commonly referred to as the "pain matrix" \cite{Tracey2007a}. (See section \ref{section:neuromatrix})

@@

 Neuropathic and acute nociceptive pain have a high degree of overlap in brain activations \cite{Moisset2007}, therefore differentiating between neuropathic and acute pain from neuroimaging continue to be an active area of research. Brain regions that are involved in the somatosensation of pain process are involved in both acute and chronic pain. These regions include the posterior insula, lateral thalamus and dorsal striatum. In chronic pain conditions however, there is increased activation in the anterior insula, ventral striatum, orbital frontal cortex and dorsal lateral prefrontal cortex \cite{Jensen2016,Alomar2016,Lorenz2005}. In chronic back pain, diffusion changes in fractional anisotropy was correlated with medial prefrontal cortex and nucleus accumbens connectivity in pain recovery, but not when the pain persistented \cite{Mansour2013}. Evidence also suggest that back pain chronification seem to increase activation in the emotional regions (amygada and medial prefrontal cortex) of the brain, and transition away from sensory region activations \cite{Hashmi2013}.
 Additionally, evidence of a "pain memory" circuitry involving the MCC \cite{Vogt2016,Nevian2017}, and abnormalities in the reward circuitry \cite{Tracey2017} have been implicated in chronic pain.

 Advances in modern machine learning methods led to their uses in attempting to improve neuropathic pain classification and diagnosis. Modern approaches combines cross-modal and additional analytical features, and can involve fMRI, diffusion imaging, as well as functional and/or structural connectivity. Machine learning in the form of multivariate pattern analysis to classify individuals with chronic back pain from normal controls, and achieve 92\% accuracy \cite{Callan2014}. Other studies in chronic pelvic pain using linear support vector machine was able to achieve 73\% accuracy \cite{Bagarinao2014}. Differentiating neuropathic and acute pain however is more difficult. Functional connectivity based classification on differentiating fibromyalgia, rheumatoid arthritis and controls was only able to achieve a diagnositc accuracy of 69\%, and classification accuracy of 78\% \cite{Sundermann2014}.
